A comparison of language control while switching within versus between languages in younger and older adults.

### 2023-08-22 manuscriptINSPIRE_R1_unmarked.Rmd
Add files via upload

---
title: Does language control differ between younger and older adults when switching
  between versus within languages?
author: "Angela de Bruin*$^1$, Heidi Kressel$^2$, & Daisy Hemmings$^2$"
output:
   #html_document: default
   bookdown::pdf_document2:
    keep_tex: true
    toc: false
    fig_caption: yes
   pdf_document: default

header-includes:
  \newcommand{\beginsupplement}{\setcounter{table}{0}       
  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} 
  \renewcommand{\thefigure}{S\arabic{figure}}}
---

@@

# flag to specify if analyses are done from scratch (takes ~6 minutes)
# if set to 0, precomputed analyses are loaded from an RData file
doanalysis <- 1

@@

#mixing cost per combi
means_LS_acc_allcombs_pivot <- means_LS_acc_allcombs_pivot %>% 
  mutate(MixingCostYoungerAction = nonswitch_ACTION_Younger - SingleNaming_ACTION_Younger,
         MixingCostOlderAction = nonswitch_ACTION_Older - SingleNaming_ACTION_Older,
         MixingCostYoungerObject = nonswitch_OBJECT_Younger - SingleNaming_OBJECT_Younger,
         MixingCostOlderObject = nonswitch_OBJECT_Older - SingleNaming_OBJECT_Older)
#get means across participants
tableS1mixswitch[2,2] <- 100*mean(means_LS_acc_allcombs_pivot$MixingCostOlderAction, na.rm=TRUE)
tableS1mixswitchSD[2,2] <- 100*sd(means_LS_acc_allcombs_pivot$MixingCostOlderAction, na.rm=TRUE)
tableS1mixswitch[1,2] <- 100*mean(means_LS_acc_allcombs_pivot$MixingCostOlderObject, na.rm=TRUE)
tableS1mixswitchSD[1,2] <- 100*sd(means_LS_acc_allcombs_pivot$MixingCostOlderObject, na.rm=TRUE)
tableS1mixswitch[2,1] <- 100*mean(means_LS_acc_allcombs_pivot$MixingCostYoungerAction, na.rm=TRUE)
tableS1mixswitchSD[2,1] <- 100*sd(means_LS_acc_allcombs_pivot$MixingCostYoungerAction, na.rm=TRUE)
tableS1mixswitch[1,1] <- 100*mean(means_LS_acc_allcombs_pivot$MixingCostYoungerObject, na.rm=TRUE)
tableS1mixswitchSD[1,1] <- 100*sd(means_LS_acc_allcombs_pivot$SwitchingCostYoungerObject, na.rm=TRUE)

@@

#mixing cost per combi
means_NLS_acc_allcombs_pivot <- means_NLS_acc_allcombs_pivot %>% 
  mutate(MixingCostYoungerPARITY = nonswitch_PARITY_Younger - SingleNaming_PARITY_Younger,
         MixingCostOlderPARITY = nonswitch_PARITY_Older - SingleNaming_PARITY_Older,
         MixingCostYoungerSIZE = nonswitch_SIZE_Younger - SingleNaming_SIZE_Younger,
         MixingCostOlderSIZE = nonswitch_SIZE_Older - SingleNaming_SIZE_Older)
#get means across participants
tableS3mixswitch[2,2] <- 100*mean(means_NLS_acc_allcombs_pivot$MixingCostOlderPARITY, na.rm=TRUE)
tableS3mixswitchSD[2,2] <- 100*sd(means_NLS_acc_allcombs_pivot$MixingCostOlderPARITY, na.rm=TRUE)
tableS3mixswitch[1,2] <- 100*mean(means_NLS_acc_allcombs_pivot$MixingCostOlderSIZE, na.rm=TRUE)
tableS3mixswitchSD[1,2] <- 100*sd(means_NLS_acc_allcombs_pivot$MixingCostOlderSIZE, na.rm=TRUE)
tableS3mixswitch[2,1] <- 100*mean(means_NLS_acc_allcombs_pivot$MixingCostYoungerPARITY, na.rm=TRUE)
tableS3mixswitchSD[2,1] <- 100*sd(means_NLS_acc_allcombs_pivot$MixingCostYoungerPARITY, na.rm=TRUE)
tableS3mixswitch[1,1] <- 100*mean(means_NLS_acc_allcombs_pivot$MixingCostYoungerSIZE, na.rm=TRUE)
tableS3mixswitchSD[1,1] <- 100*sd(means_NLS_acc_allcombs_pivot$SwitchingCostYoungerSIZE, na.rm=TRUE)

@@

#get participant mean per rule for mixing analysis
means_NV_Exp2_rule_mix <- NounVerb_Exp2_acc_outliers %>% 
  group_by(participant, Rule) %>% 
  summarise(meanRT = mean(rawrt))
head(means_NV_Exp2_rule_mix)
NV_RuleMix_RT_M <- aggregate(meanRT~Rule, data=means_NV_Exp2_rule_mix, mean, na.rm=TRUE)
NV_RuleMix_RT_SD <- aggregate(meanRT~Rule, data=means_NV_Exp2_rule_mix, sd, na.rm=TRUE)

@@

**Corresponding author:**  
Angela de Bruin  
Department of Psychology  
University of York  
York YO10 5DD  
United Kingdom

@@

While the majority of research has focused on changes in the retrieval of words, retrieval of target words alone might not be enough to allow for successful language production. We also need to manage competition from other, related words (e.g., “dog” when wanting to say “cat”). Older adults have greater difficulty suppressing irrelevant information while focusing on their goal [6]. This affects language too. For example, older adults are less coherent in story telling than younger adults, and this is modulated by their ability to suppress distracting relationships between words (e.g.,[7]).

@@

The current study examines how age influences the way people use top-down language control during language production. Furthermore, it compares whether similar control mechanisms might be used when switching between words within a language versus when switching between languages. Experiment 1 firstly examined within-language switching in a noun-verb switching task in younger and older adults. In Experiment 2, we compared this within-language switching to between-language switching in bilinguals, to examine the involvement of language control processes in both, as well as potential age-related effects.

## Bilingual language control
Bilingual language control has frequently been studied through cued language-switching paradigms, in which participants name digits or simple pictures. A cue, such as a country flag, indicates which language has to be used for a given picture. These studies typically find that bilinguals are slower to name a picture when they have to switch languages than on non-switch trials when they can use the same language as on the previous trial (switching cost; e.g.,[10];[11]). These cued switching costs reflect reactive control processes that bilinguals apply in order to respond to the language cue, activate the new target language, and avoid interference from the language used on the pervious trial (cf. [12] for a review).

@@

How much language control a bilingual needs to apply might also depend on the language they are using. A more proficient language can create more interference and bilinguals are therefore argued to apply more control over the more proficient language (L1) than over less-proficient languages (L2; [9]). Although not always observed, unbalanced bilinguals who are more proficient in their L1 than L2 have been found to show an asymmetry in switching costs (e.g., [10]), with larger costs when switching to the L1 than L2. The inhibitory control hypothesis [9] explains this in terms of time needed to overcome previously applied inhibition. When using the L2, unbalanced bilinguals might strongly suppress the L1. When returning to the L1 on the next trial, time is needed to overcome the previously applied inhibition, leading to larger switching costs. Similar asymmetries have also been observed in terms of mixing costs (e.g., [14]; [15]), with bilinguals responding faster in their L1 than L2 in single-language blocks but not in dual-language blocks. This shows language proficiency can also influence more proactive control, potentially with bilinguals over-inhibiting their L1 in dual-language contexts to allow for easier use of the L2 in the same context. In some cases, this can even lead to reversed dominance effects, with participants responding faster in their L2 than L1 in dual-language contexts (cf. [16]).

Studies assessing language changes with age in bilinguals have, similar to studies with monolinguals, shown slower or reduced lexical retrieval (e.g., [17];[18]). However, language control also appears to change with age. Several studies (e.g., [19];[20];[21]) have found both switching and mixing costs to be larger for older than younger adults, suggesting both reactive and proactive control are affected. However, Hernandez and Kohnert [22] only observed age effects in terms of mixing costs but not switching costs. In contrast, Weissberger and colleagues [20] suggested that the age effects might be strongest for switching. Finally, Calabria et al. [23] found no age effect on language switching beyond overall slower naming. Thus, based on the current literature, it remains unclear whether age affects both reactive and proactive language control.

## Language control in between- versus within-language competition
Furthermore, an open question is whether this top-down language control is specific to resolving competition between languages or is also applied when resolving competition between words within a language. In particular when weaker relationships or words have to be retrieved, for example the less dominant meaning of an ambiguous word like bank [7], top-down control might be applied. Nevertheless, in influential models of (monolingual) word production, competition between words is often resolved through bottom-up rather than top-down processes, for example by words being selected when they exceed the activation level of other words (e.g., [24];[25]). Inhibition can be part of this (e.g., [26]), but typically in the form of lateral rather than top-down inhibition, so that word representations inhibit competitors directly.  Top-down control is often not incorporated in language-production models (see [27] for a discussion; and [28] for an example of how top-down mechanisms can be implemented).

However, research has shown top-down control might be used when controlling responses in one language too. For instance, Piai and colleagues [29] compared picture naming while ignoring distractors (e.g., picture of a dog while ignoring “cat”), colour naming while ignoring distractors in a Stroop task; and a non-linguistic task requiring object discrimination while ignoring its spatial position (Simon task). Incongruent trials in all conditions showed increased activation in the anterior cingulate cortex (ACC), an area associated with domain-general attentional control.
Previous research has therefore also compared if markers associated with language control observed in bilingual language switching are also present in within-language switching. Markers such as mixing and switching costs and reversed dominance effects are difficult to explain without any form of top-down control (cf. [16];[30]). Ivanova and Hernandez [30] asked participants to name pictures using either basic-level responses (like “shoe”) or a subordinate name (like “sneaker”). Participants named pictures in blocks following the same rule or in blocks in which they had to alternate between the two rules. Switching costs, mixing costs, and asymmetries as well as reversed dominance effects were observed, leading the authors to conclude that, similar to bilingual language control, within-language competition might also trigger top-down control.

While within- and between-language control were not directly compared by Ivanova and Hernandez, Declerck and colleagues [31] compared the two by asking French-English participants to switch languages or to switch between using formal or informal words in French. Both tasks showed significant switching costs, which correlated with each other and were comparable in size. This suggests similar control mechanisms were involved in the two types of switching. However, when manipulating the interval between the cue and stimulus, only language-switching but not register-switching costs were affected. Given that the cue-stimulus interval can be interpreted as active preparation of control processes in response to the cue, this suggests a difference in control mechanisms between language- and register-switching.

Finally, some studies have assessed within- versus cross-language interference by examining language switching in a paradigm with congruent or incongruent Stroop trials. Stroop trials require participants to name the physical colour of a word, which is either the same as the word itself (e.g., “red” in red ink, congruent) or different (e.g., “green” in red ink, incongruent). Some studies have shown larger language switching costs in incongruent Stroop contexts (e.g., [32];[33]), suggesting that managing Stroop interference and between-language interference might recruit the same cognitive resources. However, not all studies have replicated these findings (e.g., [34]).

@@

## Introduction
Thus, although there are suggestions in the literature that within-language switching might use language control too, it remains unclear whether these control processes are specific to within- or between-language competition or whether they are more general and comparable across the two. In the first Experiment, we tested a within-language switching task, with the aim of comparing this task to between-language switching in Experiment 2. We chose a task asking participants to name pictures while either using a noun to describe the object or a verb describing the action associated with it (e.g., “book” or “read”). The advantage of using this paradigm is that it easily allowed us to identify the easy-to-use rule (using nouns, similar to the L1 being the easier to retrieve language) and the more difficult rule (using verbs, similar to words in the L2 being more difficult to retrieve).

We furthermore examined how different types of control, if used within one language, might change with age. Previous research on bilingual control (as also further assessed in Experiment 2) has suggested that age can influence both reactive and proactive control processes. However, research on non-linguistic control has suggested age might affect proactive control more than reactive control. Several meta-analyses (e.g., [35];[36];[37]) have shown age effects predominantly occurring on tasks requiring participants to mix or use multiple tasks. For instance, Wasylyshyn and colleagues [37] showed clear age-group differences on task-mixing costs, suggesting older adults have increased difficulty completing two tasks in parallel, potentially also in relation to changes in working memory coordinating and maintaining multiple task rules within one context. However, the same meta-analysis showed no age-group differences in terms of task-switching costs, suggesting that more reactive control mechanisms are influenced less by age.

This leaves open the possibility that age also influences proactive control more strongly in language contexts. However, language control and executive control used in non-linguistic tasks might only overlap partially (e.g., [38];[39]). Furthermore, two studies comparing age effects on bilingual switching and task switching showed different patterns depending on the task. Calabria et al. [23] only observed age-group differences on non-linguistic switching costs but not on linguistic switching costs. Weissberger et al [20] showed clear age-group differences in both tasks but while the language-switching task showed differences in terms of both switching and mixing costs, the non-linguistic switching task only showed age-group differences on the mixing cost errors (in line with e.g., [37]).

@@

As a second aim, Experiment 1 was completed by younger and older adults. If older adults have more difficulty with proactive control, we expected larger mixing costs in that age group, while their switching costs should be larger if their reactive control is affected. Finally, to examine whether any age-related differences are specific to language control, we also compared older and younger adults on a non-linguistic switching task in which they had to sort digits according to size (smaller or larger than 5) or parity (odd/even).

@@

### Participants
Experiment 1 was completed by forty older adults (Mage = 68.65, SD = 3.82, range = 65-77) and forty younger adults (Mage = 22.28, SD = 3.49, range = 18-29). 	Participants provided informed consent at the start of the study. Ethics approval was given by the ethics committee in the Department of Psychology, University of York. All participants reported that they were native speakers of English, monolingual, and had not been diagnosed with language or reading difficulties, a neurological disorder, or cognitive impairment. They also reported having normal or corrected-to-normal vision and hearing and were not taking medication that could influence their concentration. Given that the study was conducted online, we did not use an assessment of cognitive functioning (e.g., MMSE). However, we used Prolific’s screening criteria to only invite participants without a history of cognitive impairment or dementia, which participants also confirmed in the questionnaire. Sample size was determined based on effect sizes typically observed in bilingual switching tasks, which can be observed with >95% power with this sample size.  Given that age-group differences have not been studied on noun-verb switching tasks, we could not run a power analysis for the interaction between age and language control. Therefore, we followed the recommendation [40] to use 40 participants and 40 trials per combination of conditions for sufficient power in mixed-effect analyses.

Within the group of younger adults, 15 identified as male and 25 as female. Within the group of older adults, 22 identified as male and 18 as female. In terms of highest level of education achieved, 16 younger adults and 23 older adults reported having completed a graduate degree. 
An additional 12 participants were tested but excluded because they showed low accuracy (below 70% on the noun-verb switching task) in one or both of the tasks (N= 6), reported being bilingual (N=5), or reported having a neurological disorder or cognitive impairment (N=1).

@@

Per task, we conducted one set of analyses focusing on mixing effects (difference between single-rule and non-switch dual-rule trials) and one focusing on switching effects (difference between switch and non-switch trials within the dual-rule block). Each analysis included trial type (switch/non-switch; or non-switch/single-rule), rule (noun/verb for the linguistic task and size/parity for the non-linguistic task), and age group as predictors, and accuracy and reaction times as the dependent variable.

###	Materials
For the linguistic noun-verb switching task, we selected twenty pictures from the MultiPic database [41]. We selected pictures that showed high naming agreement for the nouns. Furthermore, we ensured each picture had a concrete action associated with it that could be used when responding to the “verb” rule. Feedback was gathered from three speakers to choose pictures that only had one likely action associated with it. For instance, the picture of a chair has the specific action “sit” closely linked to it. An overview of the stimuli is provided on the OSF page: https://osf.io/rdfcs/. Verb and noun picture names were matched in terms of frequency and number of phonemes (see OSF page). All verbs were one syllable long and participants were asked to just name the infinitive, without “to”. All nouns were one or two syllables long. In the non-linguistic switching task, participants saw a digit between 1 and 4 or between 6 and 9 on the screen and responded to its size or parity with a button press.

Within each task, we used two written cues per rule to avoid a confound between cue and rule switching [42]. This way, even when the rule stayed the same (e.g., two trials in a row requiring a noun response), the cue still changed. In the linguistic noun-verb task, the words “NOUN” or “OBJECT” and “VERB” or “ACTION” were used as cues. In the non-linguistic task, the words “MAGNITUDE” or “SIZE” and “PARITY” or “EVENNESS” were used.

### Procedure
Participants were recruited through Prolific.co and completed the study on Gorilla.sc [43]. After reading the information sheet and providing consent, participants first completed a short questionnaire. This questionnaire asked them about their age, gender, and education. Additional questions were included to ensure participants met our eligibility requirements (see “Participants” section). Afterwards, participants completed the linguistic and non-linguistic switching tasks, with the order counterbalanced across participants.

@@

The data and analysis code can be found here: https://github.com/AMTdeBruin/Bilingual-switching-ageing. They are also available at https://osf.io/rdfcs/.
Verbal responses in the linguistic noun-verb switching task were recorded and scored for accuracy and naming onset time. A response was scored as correct if it was the target word or a similar alternative (e.g., “gift” instead of “present”). Responses were scored as incorrect if there was no responses, if a different word was used (e.g., “apple” instead of “strawberry”), or if a response was given in the wrong language or combining two languages. Reaction times (RTs) relative to onset of picture presentation were determined using CheckVocal ([44], using CheckFile).

@@

8.	Spivey, M. J. & Marian, V. Cross talk between native and second languages: Partial activation of an irrelevant lexicon. *Psychol. Sci.* ** 10**, 281–284 (1999).

>>>>

---
title: A comparison of language control while switching within versus between languages in younger and older adults.
author: "Angela de Bruin*$^1$, Heidi Kressel$^2$, & Daisy Hemmings$^2$"
output:
   #html_document: default
   bookdown::pdf_document2:
    keep_tex: true
    toc: false
    fig_caption: yes
   pdf_document: default

header-includes:
  \newcommand{\beginsupplement}{\setcounter{table}{0}       
  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} 
  \renewcommand{\thefigure}{S\arabic{figure}}}
editor_options: 
  chunk_output_type: inline
---

@@

# flag to specify if analyses are done from scratch (takes ~6 minutes)
# if set to 0, precomputed analyses are loaded from an RData file
doanalysis <- 0

@@

#mixing cost per combi
means_LS_acc_allcombs_pivot <- means_LS_acc_allcombs_pivot %>% 
  mutate(MixingCostYoungerAction = nonswitch_ACTION_Younger - SingleNaming_ACTION_Younger,
         MixingCostOlderAction = nonswitch_ACTION_Older - SingleNaming_ACTION_Older,
         MixingCostYoungerObject = nonswitch_OBJECT_Younger - SingleNaming_OBJECT_Younger,
         MixingCostOlderObject = nonswitch_OBJECT_Older - SingleNaming_OBJECT_Older)
#get means across participants
tableS1mixswitch[2,2] <- 100*mean(means_LS_acc_allcombs_pivot$MixingCostOlderAction, na.rm=TRUE)
tableS1mixswitchSD[2,2] <- 100*sd(means_LS_acc_allcombs_pivot$MixingCostOlderAction, na.rm=TRUE)
tableS1mixswitch[1,2] <- 100*mean(means_LS_acc_allcombs_pivot$MixingCostOlderObject, na.rm=TRUE)
tableS1mixswitchSD[1,2] <- 100*sd(means_LS_acc_allcombs_pivot$MixingCostOlderObject, na.rm=TRUE)
tableS1mixswitch[2,1] <- 100*mean(means_LS_acc_allcombs_pivot$MixingCostYoungerAction, na.rm=TRUE)
tableS1mixswitchSD[2,1] <- 100*sd(means_LS_acc_allcombs_pivot$MixingCostYoungerAction, na.rm=TRUE)
tableS1mixswitch[1,1] <- 100*mean(means_LS_acc_allcombs_pivot$MixingCostYoungerObject, na.rm=TRUE)
tableS1mixswitchSD[1,1] <- 100*sd(means_LS_acc_allcombs_pivot$MixingCostYoungerObject, na.rm=TRUE)

@@

#mixing cost per combi
means_NLS_acc_allcombs_pivot <- means_NLS_acc_allcombs_pivot %>% 
  mutate(MixingCostYoungerPARITY = nonswitch_PARITY_Younger - SingleNaming_PARITY_Younger,
         MixingCostOlderPARITY = nonswitch_PARITY_Older - SingleNaming_PARITY_Older,
         MixingCostYoungerSIZE = nonswitch_SIZE_Younger - SingleNaming_SIZE_Younger,
         MixingCostOlderSIZE = nonswitch_SIZE_Older - SingleNaming_SIZE_Older)
#get means across participants
tableS3mixswitch[2,2] <- 100*mean(means_NLS_acc_allcombs_pivot$MixingCostOlderPARITY, na.rm=TRUE)
tableS3mixswitchSD[2,2] <- 100*sd(means_NLS_acc_allcombs_pivot$MixingCostOlderPARITY, na.rm=TRUE)
tableS3mixswitch[1,2] <- 100*mean(means_NLS_acc_allcombs_pivot$MixingCostOlderSIZE, na.rm=TRUE)
tableS3mixswitchSD[1,2] <- 100*sd(means_NLS_acc_allcombs_pivot$MixingCostOlderSIZE, na.rm=TRUE)
tableS3mixswitch[2,1] <- 100*mean(means_NLS_acc_allcombs_pivot$MixingCostYoungerPARITY, na.rm=TRUE)
tableS3mixswitchSD[2,1] <- 100*sd(means_NLS_acc_allcombs_pivot$MixingCostYoungerPARITY, na.rm=TRUE)
tableS3mixswitch[1,1] <- 100*mean(means_NLS_acc_allcombs_pivot$MixingCostYoungerSIZE, na.rm=TRUE)
tableS3mixswitchSD[1,1] <- 100*sd(means_NLS_acc_allcombs_pivot$MixingCostYoungerSIZE, na.rm=TRUE)

@@

#get participant mean per rule for mixing analysis
means_NV_Exp2_rule_mix <- NounVerb_Exp2_acc_outliers_mixing %>% 
  group_by(participant, Rule) %>% 
  summarise(meanRT = mean(rawrt))
head(means_NV_Exp2_rule_mix)
NV_RuleMix_RT_M <- aggregate(meanRT~Rule, data=means_NV_Exp2_rule_mix, mean, na.rm=TRUE)
NV_RuleMix_RT_SD <- aggregate(meanRT~Rule, data=means_NV_Exp2_rule_mix, sd, na.rm=TRUE)

@@

**Corresponding author:**  
Angela de Bruin  
Department of Psychology  
University of York  
York YO10 5DD  
United Kingdom  
angela.debruin@york.ac.uk

@@

While the majority of research has focused on changes in the retrieval of words, retrieval of target words alone might not be enough to allow for successful language production. We also need to manage competition from other, related words (e.g., “dog” when wanting to say “cat”). Older adults have greater difficulty suppressing irrelevant information while focusing on their goal [6]. This affects language too. For example, older adults are less coherent in story telling than younger adults, and this is modulated by their ability to suppress distracting relationships between words (e.g., [7]).

@@

The current study examines how age influences the way people use language control during language production. Furthermore, it compares whether similar control mechanisms might be used when switching between words within a language versus when switching between languages. Experiment 1 firstly examined within-language switching in a noun-verb switching task in younger and older adults. In Experiment 2, we compared this within-language switching to between-language switching in bilinguals, to examine the involvement of language control processes in both, as well as potential age-related effects.

## Bilingual language control
Bilingual language control has frequently been studied through cued language-switching paradigms, in which participants name digits or simple pictures. A cue, such as a country flag, indicates which language has to be used for a given picture. These studies typically find that bilinguals are slower to name a picture when they have to switch languages than on non-switch trials when they can use the same language as on the previous trial (switching cost; e.g., [10];[11]). These cued switching costs reflect reactive control processes that bilinguals apply in order to respond to the language cue, activate the new target language, and avoid interference from the language used on the previous trial (cf. [12] for a review).

@@

Prominent models of language control (e.g., [9]) focus on language control in the form of inhibition, with bilinguals inhibiting the non-target language. How much inhibition is applied might depend on the language they are using. A more proficient language can create more interference and bilinguals are therefore argued to apply more control over the more proficient language (L1) than over less-proficient languages (L2; [9]). Although not always observed, unbalanced bilinguals who are more proficient in their L1 than L2 have been found to show an asymmetry in switching costs (e.g., [10]), with larger costs when switching to the L1 than L2. The inhibitory control hypothesis [9] explains this in terms of time needed to overcome previously applied inhibition. When using the L2, unbalanced bilinguals might strongly suppress the L1. When returning to the L1 on the next trial, time is needed to overcome the previously applied inhibition, leading to larger switching costs. Similar asymmetries have also been observed in terms of mixing costs (e.g., [14]; [15]), with bilinguals responding faster in their L1 than L2 in single-language blocks but not in dual-language blocks. This shows language proficiency can also influence more proactive control, potentially with bilinguals over-inhibiting their L1 in dual-language contexts to allow for easier use of the L2 in the same context. In some cases, this can even lead to reversed dominance effects, with participants responding faster in their L2 than L1 in dual-language contexts (cf. [16]). However, these findings can also be explained without purely relying on inhibition. Instead (or in addition), bilinguals might over-activate their L2 to allow for easier use in a dual-language context [17]. Following these accounts, (relatively) faster L2 than L1 responses in a dual-language context might be the consequence of this L2 over-activation.

Studies assessing language changes with age in bilinguals have, similar to studies with monolinguals, shown slower or reduced lexical retrieval (e.g., [18];[19]). However, language control also appears to change with age. Several studies (e.g., [20];[21];[22]) have found both switching and mixing costs to be larger for older than younger adults, suggesting both reactive and proactive control are affected. However, Hernandez and Kohnert [23] only observed age-group differences in terms of mixing costs but not switching costs. In contrast, Weissberger and colleagues [21] suggested that the age-group differences might be strongest for switching. Finally, Calabria et al. [24] found no age-group differences on language switching beyond overall slower naming. Thus, based on the current literature, it remains unclear whether age affects both reactive and proactive language control.

## Language control in between- versus within-language competition
Furthermore, an open question is whether this language control is specific to resolving competition between languages or is also applied when resolving competition between words within a language. In particular when weaker relationships or words have to be retrieved, for example the less dominant meaning of an ambiguous word like bank [7], top-down control might be applied. In bilingual language production models (e.g., [9]), language control involves both bottom-up processes (e.g., an external cue telling the bilingual which language to use) and top-down control processes (e.g., inhibition and/or activation of specific languages). However, in influential models of (monolingual) word production, competition between words is often resolved just through bottom-up processes rather than top-down processes, for example by words being selected when they exceed the activation level of other words (e.g., [25];[26]). Inhibition can be part of this (e.g., [27]), but typically in the form of lateral rather than top-down inhibition, so that word representations inhibit competitors directly.  Top-down control is often not incorporated in these models (see [28] for a discussion; and [29] for an example of how top-down mechanisms can be implemented).

However, research has shown top-down control might be used when controlling responses in one language too. For instance, Piai and colleagues [30] compared picture naming while ignoring distractors (e.g., picture of a dog while ignoring “cat”), colour naming while ignoring distractors in a Stroop task; and a non-linguistic task requiring object discrimination while ignoring its spatial position (Simon task). Incongruent trials in all conditions showed increased activation in the anterior cingulate cortex (ACC), an area associated with domain-general attentional control.
Previous research has therefore also compared if markers associated with language control observed in bilingual language switching are also present in within-language switching. Markers such as mixing and switching costs and reversed dominance effects are difficult to explain without any form of top-down control (cf. [16];[31]). Ivanova and Hernandez [31] asked participants to name pictures using either basic-level responses (like “shoe”) or a subordinate name (like “sneaker”). Participants named pictures in blocks following the same rule or in blocks in which they had to alternate between the two rules. Switching costs, mixing costs, and asymmetries as well as reversed dominance effects were observed, leading the authors to conclude that, similar to bilingual language control, within-language competition might also trigger top-down control.

While within- and between-language control were not directly compared by Ivanova and Hernandez, Declerck and colleagues [32] compared the two by asking French-English participants to switch languages or to switch between using formal or informal words in French. Both tasks showed significant switching costs, which correlated with each other and were comparable in size. This suggests similar control mechanisms were involved in the two types of switching. However, when manipulating the interval between the cue and stimulus, only language-switching but not register-switching costs were affected. Given that the cue-stimulus interval can be interpreted as active preparation of control processes in response to the cue, this suggests a difference in control mechanisms between language- and register-switching. Cattaneo and colleagues [33] also observed some similarities and some differences when comparing between-language switching to a noun-verb within-language switching task. Switching costs did not correlate between the two tasks. In contrast, mixing costs did correlate between the tasks, although this appeared to be driven by the group of Parkinson's Disease patients rather than the control group of older adults.

Finally, some studies have assessed within- versus cross-language interference by examining language switching in a paradigm with congruent or incongruent Stroop trials. Stroop trials require participants to name the physical colour of a word, which is either the same as the word itself (e.g., “red” in red ink, congruent) or different (e.g., “green” in red ink, incongruent). Some studies have shown larger language-switching costs in incongruent Stroop contexts (e.g., [34];[35]), suggesting that managing Stroop interference and between-language interference might recruit the same cognitive resources. However, not all studies have replicated these findings (e.g., [36]).

@@

## Introduction
Thus, although there are suggestions in the literature that within-language switching might use language control too, it remains unclear whether these control processes are specific to within- or between-language competition or whether they are more general and comparable across the two. In the first Experiment, we tested a within-language switching task, with the aim of comparing this task to between-language switching in Experiment 2. We chose a task (similar to [33]) asking participants to name pictures while either using a noun to describe the object or a verb describing the action associated with it (e.g., “book” or “read”). The advantage of using this paradigm is that previous research has shown nouns are processed faster than verbs (e.g., [33];[37];[38], cf. [39] for a review including lesion research), allowing us to identify a priori the easy-to-use rule (using nouns, similar to the L1 being the easier to retrieve language) and the more difficult rule (using verbs, similar to words in the L2 being more difficult to retrieve).

We furthermore examined how different types of control, if used within one language, might change with age. Previous research on bilingual control (as also further assessed in Experiment 2) has suggested that age can influence both reactive and proactive control processes. However, research on non-linguistic control has suggested age might affect proactive control more than reactive control. Several meta-analyses (e.g., [40];[41];[42]) have shown age effects predominantly occurring on tasks requiring participants to mix or use multiple tasks. For instance, Wasylyshyn and colleagues [42] showed clear age-group differences on task-mixing costs, suggesting older adults have increased difficulty completing two tasks in parallel, potentially also in relation to changes in working memory coordinating and maintaining multiple task rules within one context. However, the same meta-analysis showed no age-group differences in terms of task-switching costs, suggesting that more reactive control mechanisms are influenced less by age.

This leaves open the possibility that age also influences proactive control more strongly in language contexts. However, language control and executive control used in non-linguistic tasks might only overlap partially (e.g., [43];[44]). Furthermore, two studies comparing age effects on bilingual switching and task switching showed different patterns depending on the task. Calabria et al. [24] only observed age-group differences on non-linguistic switching costs but not on linguistic switching costs. Weissberger et al [21] showed clear age-group differences in both tasks but while the language-switching task showed differences in terms of both switching and mixing costs, the non-linguistic switching task only showed age-group differences on the mixing cost errors (in line with e.g., [42]).

@@

As a second aim, Experiment 1 was completed by younger and older adults. If older adults have more difficulty with proactive control, we expected larger mixing costs in that age group, while their switching costs should be larger if their reactive control is affected. Finally, to examine whether any age-related differences are specific to language control, we also compared older and younger adults on a non-linguistic switching task in which they had to sort digits according to size (smaller or larger than 5) or parity (odd/even). For the non-linguistic switching task, we expected (in line with meta-analyses such as [42]) proactive control (mixing costs) to differ between age group while reactive control (switching costs) might be not or less affected.

@@

### Participants
Experiment 1 was completed by forty older adults (*M*age = 68.65 years old, *SD* = 3.82, range = 65-77) and forty younger adults (*M*age = 22.28, *SD* = 3.49, range = 18-29). 	Participants provided informed consent at the start of the study. Ethics approval was given by the ethics committee in the Department of Psychology, University of York. The study was conducted in accordance with the Declaration of Helsinki, with the exception of the requirement for pre-registration. All participants reported that they were native speakers of English, monolingual, and had not been diagnosed with language or reading difficulties, a neurological disorder, or cognitive impairment. They also reported having normal or corrected-to-normal vision and hearing and were not taking medication that could influence their concentration. Given that the study was conducted online, we did not use an assessment of cognitive functioning (e.g., MMSE). However, we used Prolific’s screening criteria to only invite participants without a history of cognitive impairment or dementia, which participants also confirmed in the questionnaire. Sample size was determined based on effect sizes typically observed for switching and mixing effects in bilingual switching tasks, which can be observed with >95% power with this sample size.  Given that age-group differences have not been studied on noun-verb switching tasks, we could not run a power analysis for the interaction between age and language control. Therefore, we followed the recommendation [45] to use 40 participants and 40 trials per combination of conditions for sufficient power in mixed-effect analyses.

Within the group of younger adults, 15 identified as male and 25 as female. Within the group of older adults, 22 identified as male and 18 as female. In terms of highest level of education achieved, 16 younger adults and 23 older adults reported having completed a graduate degree. 
An additional 12 participants were tested but excluded because they showed low accuracy (below 70% on the noun-verb switching task; N= 6), reported being bilingual (N=5), or reported having a neurological disorder or cognitive impairment (N=1).

@@

Per task, we conducted one set of analyses focusing on mixing effects (difference between single-rule and non-switch dual-rule trials) and one focusing on switching effects (difference between switch and non-switch trials within the dual-rule block). Each analysis included trial type (switch/non-switch; or non-switch/single-rule), rule (noun/verb for the linguistic task and size/parity for the non-linguistic task), and age group as predictors, and accuracy and reaction times as the dependent variables.

###	Materials
For the linguistic noun-verb switching task, we selected twenty pictures from the MultiPic database [46]. We selected pictures that showed high naming agreement for the nouns. Furthermore, we ensured each picture had a concrete action associated with it that could be used when responding to the “verb” rule. Feedback was gathered from three speakers to choose pictures that only had one likely action associated with it. For instance, the picture of a chair has the specific action “sit” closely linked to it. An overview of the stimuli is provided on the OSF page: https://osf.io/rdfcs/. Verb and noun picture names were matched in terms of frequency and number of phonemes (see OSF page). All verbs were one syllable long and participants were asked to just name the infinitive, without “to”. All nouns were one or two syllables long. In the non-linguistic switching task, participants saw a digit between 1 and 4 or between 6 and 9 on the screen and responded to its size or parity with a button press.

Within each task, we used two written cues per rule to avoid a confound between cue and rule switching [47]. This way, even when the rule stayed the same (e.g., two trials in a row requiring a noun response), the cue still changed. In the linguistic noun-verb task, the words “NOUN” or “OBJECT” and “VERB” or “ACTION” were used as cues. In the non-linguistic task, the words “MAGNITUDE” or “SIZE” and “PARITY” or “EVENNESS” were used.

### Procedure
Participants were recruited through Prolific.co and completed the study on Gorilla.sc [48]. After reading the information sheet and providing consent, participants first completed a short questionnaire. This questionnaire asked them about their age, gender, and education. Additional questions were included to ensure participants met our eligibility requirements (see “Participants” section). Afterwards, participants completed the linguistic and non-linguistic switching tasks, with the order counterbalanced across participants.

@@

The data and analysis code can be found here: https://github.com/AMTdeBruin/Bilingual-switching-ageing. They are also available at https://osf.io/rdfcs/.
Verbal responses in the linguistic noun-verb switching task were recorded and scored for accuracy and naming onset time. A response was scored as correct if it was the target word or a similar alternative (e.g., “gift” instead of “present”). Responses were scored as incorrect if there was no response, if a different word was used (e.g., “apple” instead of “strawberry”), or if a response was given in the wrong language or combining two languages. Reaction times (RTs) relative to onset of picture presentation were determined using CheckVocal ([49], using CheckFile).

@@

8.	Spivey, M. J. & Marian, V. Cross talk between native and second languages: Partial activation of an irrelevant lexicon. *Psychol. Sci.* **10**, 281–284 (1999).
